---
title: "Clustering Examples with Penguins"
output: html_notebook
---

# Load Required Libraries
```{r}
library(tidyverse)
library(tidymodels)
library(GGally)
library(factoextra)
set.seed(400)
```

# Read Data
```{r}
library(palmerpenguins)
penguins
```

# Explore Data
```{r}
penguins %>%
  ggpairs()

ggsave(
  filename = "pairs.png",
  device = png,
  path = "../plots",
  width = 28,
  height = 16,
  units = "in"
)
```

# K Means Clustering
*This tutorial is heavily based on https://juliasilge.com/blog/kmeans-employment/

## Preprocess (Bake) Data
```{r}
clustering_rec <- recipe(~., data = penguins) %>%
  step_impute_mean(all_numeric()) %>%
  step_normalize(all_numeric())
```


```{r}
clustering_rec %>%
  prep() %>%
  tidy(2)
```


```{r}
baked_penguins <- clustering_rec %>%
  prep() %>%
  bake(penguins) %>%
  select(is.numeric, -year)

baked_penguins
```

## Apply Clustering
```{r}
kclust <- baked_penguins %>% kmeans(centers = 6)
kclust
```

## Check Number of Cluster
```{r}
k_params = 2:20
tot.withinss <- c()

for(k in k_params){
  kclust_temp <- baked_penguins %>% kmeans(centers = k)
  tot.withinss <- tot.withinss %>% 
    append(kclust_temp$tot.withinss)
}

elbow_values <- tibble(
  k_params,
  tot.withinss
)
```

```{r}
elbow_values %>%
  ggplot(aes(x=k_params, y=tot.withinss)) +
  geom_line()
```

## Investigate Clusters
```{r}
augment(kclust, penguins)
```

# Hierarchical Clustering
## Preprocess (Bake) Data
```{r}
clustering_rec <- recipe(~., data = penguins) %>%
  step_impute_mean(all_numeric()) %>%
  step_normalize(all_numeric())
```

```{r}
baked_penguins <- clustering_rec %>%
  prep() %>%
  bake(penguins) %>%
  select(is.numeric, -year)

baked_penguins
```

## Apply Clustering
```{r}
hier_clust <- baked_penguins %>%
  dist() %>%
  hclust()
```

## Investigate Clusters
```{r}
hier_clust %>%
  fviz_dend(k=6)
```

```{r}
penguins %>%
  mutate(.cluster = as.factor(hier_clust %>% cutree(k = 3))) %>%
  ggpairs(columns = 1:8,
          aes(colour = .cluster))

ggsave(
  filename = "hierarchical_pairs.png",
  device = png,
  path = "../plots",
  width = 28,
  height = 16,
  units = "in"
)
```

For additional tips on investigating results for Hierarchical Clustering, check out https://juliasilge.com/blog/health-care-indicators/